---
title: "3 Functions and Macros Homework"
author: "Li Zhang"
date: "06/17/2021"
output:
  pdf_document: default
  html_document: default
  word_document: default
version: 1.2
---


# Exercise 1

As before, using values from Table 1, show how Wansink and Payne calculate percent increase in calories per recipe (43.7%), calories per serving (63.0%) and servings per recipe (-2.0). That is, confirm the values in the final column of Table 1 (Change from
1936 to 2006, %, p. 292).

Write a function to perform these calculations. Name the function `percent.increase` (or similar for SAS) with parameter `m_1` and `m_2`. Check that this function duplicates the values you calculated for Homework 2.

## Answer

```{r}

#Homework2:
scales::percent((3051.9-2123.8)/2123.8)

scales::percent((384.4-268.1)/268.1)

scales::percent((12.7-12.9)/12.9)



# Homework3:Write a function to perform these calculations:
percent.increase<-function(m_1, m_2){
  return(scales::percent((m_1-m_2)/m_2))
}
percent.increase(3051.9, 2123.8)
percent.increase(384.4, 268.1)
percent.increase(12.7,12.9)

#PMC: The answers are the same as those I calculated for Homework2
```

# Exercise 2

Consider this quotation from Wansink

>Over the past 70 years, the total caloric content increased
for 14 of the 18 recipes. Because of changes in ingredients,
the mean average calories in a recipe increased by 928.1 (from
2123.8 calories [95% CI, 1638.7 to 2608.9 calories] to 3051.9 calories
[CI, 2360.7 to 3743.1 calories]), representing a 43.7% increase
(P < 0.001). As the Table indicates, mean average calories per serving
increased for 17 of 18 recipes and was influenced by both
changes in ingredients and changes in serving size. The resulting
increase of 168.8 calories (from 268.1 calories [CI, 210.4 to 325.8
calories] to 436.9 calories [CI, 359.1 to 514.7 calories]) represents a
63.0% increase (P < 0.001) in calories per serving.

Write a function named `ConfidenceInterval` that returns list comprised of the `Lower` and `Upper` bounds. 
You may `ConfidenceBound` function from course outline to produce confidence interval or write your own version. You function should accept as arguments `mean`, `sd` and `n`, and an optional parameter `alpha=0.05`.

Confirm the confidence intervals reported as

- [95% CI, 1638.7 to 2608.9 calories]
- [CI, 2360.7 to 3743.1 calories]
- [CI, 210.4 to 325.8 calories]
- [CI, 359.1 to 514.7 calories]

Use the appropriate values from Table 1. 

Then, write a function `ConfidenceBoundT` that uses the $t$ distribution instead of the normal distribution (call `pt` instead of `pnorm` and use $n-1$ d.f.). Compare the resulting confidence intervals.

If you use SAS, define a subroutine in IML and accepts Upper and Lower as parameters, then print these values after each function call. If you wish, you may return a matrix (even though we haven't covered them in lecture).

## Answer

Define your function(s) in the code chunk below, the call the function with appropriate parameters in the following sections

```{r}

StandardError <- function(sd,n){
  sd/sqrt(n)
}

ConfidenceBound<-function(sd, n, alpha=0.05){
  qnorm(1-alpha/2)*StandardError(sd,n)
}

ConfidenceInterval<-function(mean, sd, n, alpha=0.05){
  local.Lower<- mean-ConfidenceBound(sd, n)
  local.Upper<- mean+ConfidenceBound(sd, n)
  return(list(Lower=local.Lower, Upper=local.Upper)) 
}

ConfidenceBoundT <- function(sd, n, alpha = 0.05) {
qt(1-alpha/2, n-1)*StandardError(sd,n)
}

ConfidenceIntervalT<-function(mean, sd, n, alpha=0.05){
  local.Lower<- mean-ConfidenceBoundT(sd, n)
  local.Upper<- mean+ConfidenceBoundT(sd, n)
  return(list(LowerT=local.Lower, UpperT=local.Upper)) 
}



```

## [95% CI, 1638.7 to 2608.9 calories]


```{r}
StandardError(1050.0,18)
ConfidenceBound(1050.0, 18)
ConfidenceInterval(2123.8,1050.0, 18)
ConfidenceBoundT(1050.0, 18)
ConfidenceIntervalT(2123.8,1050.0, 18)
# when use normal distribution to calculate connfidence intervals, 
# the results are the same as the ones reported by Wansink.[1638.7, 2608.9].
# ConfidenceBoundT(t-distribution)>ConfidenceBound(normal distribution)
# ConfidenceIntervalT is wider than ConfidenceInterval.
```

## [CI, 2360.7 to 3743.1 calories]

```{r}

StandardError(1496.2,18)
ConfidenceBound(1496.2, 18)
ConfidenceInterval(3051.9,1496.2, 18)
ConfidenceBoundT(1496.2, 18)
ConfidenceIntervalT(3051.9,1496.2, 18)
# when use normal distribution to calculate connfidence intervals, 
# the results are the same as the ones reported by Wansink.[2360.7, 3743.1].
# ConfidenceBoundT(t-distribution)>ConfidenceBound(normal distribution)
# ConfidenceIntervalT is wider than ConfidenceInterval.
```

## [CI, 210.4 to 325.8 calories]

```{r}

StandardError(124.8,18)
ConfidenceBound(124.8, 18)
ConfidenceInterval(268.1,124.8, 18)
ConfidenceBoundT(124.8, 18)
ConfidenceIntervalT(268.1,124.8, 18)
# when I use normal distribution to calculate connfidence intervals, 
# the results are the same as the ones reported by Wansink.[210.4, 325.8].
# ConfidenceBoundT(t-distribution)>ConfidenceBound(normal distribution)
# ConfidenceIntervalT is wider than ConfidenceInterval.
```

## [CI, 359.1 to 514.7 calories]

```{r}

StandardError(168.3,18)
ConfidenceBound(168.3, 18)
ConfidenceInterval(384.4,168.2, 18)
ConfidenceBoundT(168.3, 18)
ConfidenceIntervalT(384.4,168.2, 18)

#Bases on Table1, Mean Average calories per serving on 2006 should be 384.4 instead of 436.9, 
#since 436.9 is F value. 
#Therefore, the confidence intervals I calculated here [306.697, 462.103] 
#is different from the one reported by Wansink [359.1, 514.7].
#ConfidenceBoundT(t-distribution)>ConfidenceBound(normal distribution)
#ConfidenceIntervalT is wider than ConfidenceInterval.



##Brief Comment: Based on my calculation, I found that 
#ConfidenceBoundT(use normal distribution) is greater than ConfidenceBound(use t-distribution), 
# and naturally lead that the confidence Intervals calculated by using t-distribution 
# is wider than the confidence intervals reported by Wansink(using normal distribution).
#As we know,t-distribution is more useful when the sample size is small(less than 30) 
#while normal distribution is for larger sample size(greater than 30). 
#The basic idea here is that 
#when we have small sample sizes(the sample size n is 18 in our case),
#we're less certain about the true population mean. 
#so it makes sense to use the t-distribution to produce wider confidence intervals
#that have a higher chance of containing the true population mean.
```


# Exercise 3

Implement the probability density function for the normal distribution as a function or macro. 

$$
f(x,  \mu, \sigma^2) = \frac{1}{\sigma \sqrt{2 \pi}} e^{- \frac{(x - \mu)^2}{2 \sigma^2}}
$$

Define $\mu$ and $\sigma$ as optional parameters, taking values `mu=0` and `sigma=1`. Name this function `norm.pdf`
  
## Answer

Define your function(s) in the code chunk below, the call the function with appropriate parameters in the following sections. Use $\sigma = 2$ and compare with previous homework.

```{r}
norm.pdf<-function(x, mu=0, sigma=1){
m<- exp(-((x-mu)^2)/(2*(sigma)^2))
n<- 1/(sigma*(sqrt(2*pi)))
return(m*n)
}

```

## $x=-0.1$

```{r}
norm.pdf(-0.1, 0, 2)
```

## $x=0.0$

```{r}
norm.pdf(0.0, 0, 2)
```

## $x=0.1$

```{r}
norm.pdf(0.1, 0, 2)

#PMC:The answers are the same as the results of the previous homework. 
#Here we need to know that 
#If the caller provides an arguments for any one of a succession of optional parameters, 
#it must provide arguments for all preceding optional parameters.
```


# Exercise 4

The probability mass function for value $x$ from Poisson data with a mean and variance $\lambda$ is given by 

$$
f(x,\lambda) = \frac{e^{-\lambda} \lambda^x}{x!}
$$
Write a function `pois.pmf` that accepts two parameters, `x` and `lambda`. Use the built in `factorial` function for $x!$. Note that $x$ should be an integer value, so call a rounding function inside your function.

Test your function with $\lambda = 12$ at $x = 4,12,20$

## Answer

Define your function(s) in the code chunk below, the call the function with appropriate parameters in the following sections

```{r}
pois.pmf<-function(x,lambda){
  ((exp(-lambda))*(lambda)^round(x))/factorial(round(x))
}
```

## $x=4$

```{r}
pois.pmf(4,12)
```

## $x=12$

```{r}
pois.pmf(12,12)
```

## $x=20$

```{r}
pois.pmf(20,12)
```

You can check your work against the built in Poisson distribution functions.

```{r}
dpois(4,12)
dpois(12,12)
dpois(20,12)
# PMC: we know that possion distribution is not defined for non-intergers, 
# therefore, when we write our own function: pois.pmf
# we need to  make the decision about how to handle non-interger input. 
#In my self defined function pois.pmf, 
#I used rounding function to round every parameter x. 
#The return value would be the same as the result of bulit-in the function dpois.
#when lambda and x are the same.
```




