---
title: "2 Calculation Exercises"
author: 'Li Zhang'
date: '06/11/2021'
output:
  pdf_document: default
  word_document: default
  html_document: default
version: 1.2
---

# General instructions.

There are four exercises below, and you will be required to provide four solutions, each worth ten points. You may provide a solution in either R or SAS, at your discretion. You must submit results document in Word or PDF form, along with your original Rmd (no R scripts) or SAS code for full credit. However, answers to any discussion questions must be included in the PDF or Word file submitted to D2L; we will not be grading SAS code files.


# Exercise 1

Using values from Table, show how Wansink and Payne calculate percent increase in calories per recipe (43.7%), calories per serving (63.0%) and servings per recipe (-2.0). That is, confirm the values in the final column of Table 1 (Change from
1936 to 2006, %, p. 292).

## Answer

```{r}

print((3051.9-2123.8)/2123.8)

print((384.4-268.1)/268.1)

print((12.7-12.9)/12.9)

```

# Exercise 2

Using the values from Wansink, Table 1 as given in Homework 1 or in the course outline, calculate $p$-values for the difference between calories per serving for each of the pairs of means listed below:

- 1936 versus 2006
- 1936 versus 1997
- 1997 versus 2006

You may use either Wald's test or Student's $t$.

How do your calculated $p$-values compare to the F test $p$-value reported in Table 1. Does this test provide stronger or weaker evidence for the claim that calories per serving has changed over time?

## Answer

Enter the R code in the chunks below. If you choose SAS for this exercise, use the marked portion in the SAS homework template. 
## Null Hypothesis: calories per serving hasn't changed over time.

```{r}
#### 1936 versus 2006
m_1936 <-268.1
m_2006 <-384.4
s_1936 <-124.8
s_2006 <-168.3
n_1936 <-n_2006 <-18
hat.se <- sqrt(((s_1936^2)/n_1936)+((s_2006^2)/n_2006))
w<-(m_2006-m_1936)/hat.se
t<-(m_2006-m_1936)/hat.se
2*(1-pt(abs(t),(n_1936-1)+ n_2006-1))
# Null Hypothesis:calories per serving hasn't changed from 1936 to 2006
# p<=0.05, reject null hypothesis. 
# We conclude that calories per serving has changed from 1936 to 2006. 
# This test provide the stronger evidence that calories per serving has changed over time.
```



```{r}
#### 1936 versus 1997
m_1936 <-268.1
m_1997 <-288.6
s_1936 <-124.8
s_1997 <-122.0
n_1936 <-n_1997 <-18
hat.se <- sqrt(((s_1936^2)/n_1936)+((s_1997^2)/n_1997))
w<-(m_1997-m_1936)/hat.se
t<-(m_1997-m_1936)/hat.se
2*(1-pt(abs(t),(n_1936-1)+ n_1997-1))
# Null Hypothesis:calories per serving hasn't changed from 1936 to 1997
# p>0.05, fail to reject null hypothesis. 
# We conclude that calories per serving hasn't changed from 1936 to 1997. 
# This test provide the weaker evidence that calories per serving has changed over time.
```


```{r}
#### 1997 versus 2006
m_1997 <-288.6
m_2006 <-384.4
s_1997 <-122.0
s_2006 <-168.3
n_1997 <-n_2006 <-18
hat.se <- sqrt(((s_1997^2)/n_1997)+((s_2006^2)/n_2006))
w<-(m_2006-m_1997)/hat.se
t<-(m_2006-m_1997)/hat.se
2*(1-pt(abs(t),(n_1997-1)+ n_2006-1))

# Null Hypothesis:calories per serving hasn't changed from 1997 to 2006
# p>0.05, fail to reject null hypothesis. 
# We conclude that calories per serving hasn't changed from 1997 to 2006. 
# This test provide the weaker evidence that calories per serving has changed over time.
```

# Exercise 3

Using the values from Wansink, Table 1 as given in Homework 1 or in the course outline, calculate $p$-values for the difference between servings per recipe for each of the pairs of means listed below:

- 1946 and 1951
- 1951 and 1963
- 1997 and 2006

You may use either Wald's test or Student's $t$.

Do your calculate $p$-values support this statement from Wansink (p.291)?

> Over the 70-year history of The Joy of Cooking, the recommended serving sizes were altered at 3 points. Between 1946 and 1951, 3 of 18 recipes increased their serving size by an average of 32.5%. Between 1951 and 1963, 4 recipes increased their serving size by an average of 20.0%. Between 1997 and 2006, 5 recipes increased their serving size by an average of 21.1%

```{r}
#### 1946 and 1951
m_1946 <-12.9
m_1951 <-13.0
s_1946 <-13.3
s_1951 <-14.5
n_1946 <-n_1951 <-18
hat.se <- sqrt(((s_1946^2)/n_1946)+((s_1951^2)/n_1951))
w<-(m_1951-m_1946)/hat.se
t<-(m_1951-m_1946)/hat.se
2*(1-pt(abs(t), (n_1946-1 +n_1951-1)))
# Null Hypothesis:serving per recipe hasn't changed from 1946 to 1951
# p> 0.05. fail to reject the null hypothesis. 
# We conclude that the difference between servings per recipe from 1946 to 1951 is 0; or
# We conclude that servings per recipe hasn't changed from 1946 to 1951. 
# Therefore, our calculate cannot support the statement that 
# Between 1946 and 1951, 3 of 18 recipes increased their serving size by an average of 32.5%.

#### 1951 and 1963
m_1951 <-13.0
m_1963 <-12.7
s_1951 <-14.5
s_1963 <-14.6
n_1951 <-n_1963 <-18
hat.se <- sqrt(((s_1951^2)/n_1951)+((s_1963^2)/n_1963))
w<-(m_1963-m_1951)/hat.se
t<-(m_1963-m_1951)/hat.se
2*(1-pt(abs(t), (n_1951-1 +n_1963-1)))
# Null Hypothesis:serving per recipe hasn't changed from 1951 to 1963
# p> 0.05. fail to reject the null hypothesis. 
# We conclude that the difference between servings per recipe from 1951 to 1963 is 0; or
# We conclude that servings per recipe hasn't changed from 1951 to 1963. 
# Therefore, our calculate cannot support the statement that 
# Between 1951 and 1963, 4 recipes increased their serving size by an average of 20.0%.

#### 1997 and 2006
m_1997 <-12.4
m_2006 <-12.7
s_1997 <-14.3
s_2006 <-13.0
n_1997 <-n_2006 <-18
hat.se <- sqrt(((s_1997^2)/n_1997)+((s_2006^2)/n_2006))
w<-(m_2006-m_1997)/hat.se
t<-(m_2006-m_1997)/hat.se
2*(1-pt(abs(t), (n_1997-1 +n_2006-1)))
# Null Hypothesis:serving per recipe hasn't changed from 1997 to 2006
# p> 0.05. fail to reject the null hypothesis. 
# We conclude that the difference between servings per recipe from 1997 to 2006 is 0; or
# We conclude that servings per recipe hasn't changed from 1997 to 2006. 
# Therefore, our calculate cannot support the statement that 
# Between 1997 and 2006, 5 recipes increased their serving size by an average of 21.1%.
```
# Exercise 4

The probability density function for the distribution of a random variable denoted $x$, when taken from a normal population with mean $\mu$ and variance $\sigma^2$ is calculated by

$$
f(x,  \mu, \sigma^2) = \frac{1}{\sigma \sqrt{2 \pi}} e^{- \frac{(x - \mu)^2}{2 \sigma^2}}
$$
For values of $x = \{-0.1, 0.0, 0.1 \}$, write code to implement the formula above, for $\mu = 0$ and $\sigma = 2$. 
  
## Answer

Enter the R code in the chunks below. If you choose SAS for this exercise, use the marked portion in the SAS homework template. 

#### $x=-0.1$

```{r}
x<- -0.1
mu <- 0
sigma<-2
m=exp(-((x-mu)^2)/(2*(sigma)^2))
n=1/(sigma*(sqrt(2*pi)))
m*n
```

#### $x=0.0$

```{r}
x<-0.0
mu <-0
sigma<-2
m=exp(-((x-mu)^2)/(2*(sigma)^2))
n=1/(sigma*(sqrt(2*pi)))
m*n
```

#### $x=0.1$

```{r}
x<- 0.1
mu <-0
sigma<-2
m=exp(-((x-mu)^2)/(2*(sigma)^2))
n=1/(sigma*(sqrt(2*pi)))
m*n

```

You can confirm your results using the built in normal distribution function. Look up `dnorm` in R help and use the same values for `x`, `mean` and `sigma` as above. You should get matching results to at least 12 decimal places.


```{r}
x <- -0.1
dnorm(x, mean = 0, sd = 2, log = FALSE)
x <-0.0
dnorm(x, mean = 0, sd = 2, log = FALSE)
x <-0.1
dnorm(x, mean = 0, sd = 2, log = FALSE)

```


