---
title: "Homework6"
author: "Li Zhang"
date: "09/27/2021"
output:
  word_document: default
  html_document: default
  pdf_document: default
---

### Exercise 3.16

What is the class of the R built-in data set mtcars? Show the function that you used to check the class.
```{r}

# the class of R built-in data set mtcars is a data frame.
class(mtcars)


```




### Exercise 3.17

For data set mtcars, select a subset which satisfies that the cyl is 6 and mpg greater than 21.2. The subset should not include the variable carb.

```{r}

y<-mtcars
y.subset <- y[y$cyl == 6 & y$mpg > 21.2, -11]
y.subset


```
### Exercise 3.18

Use aggregate() function to calculate the median of the variables: mpg, disp, hp, and wt across the number of cylinders cyl in the mtcars data set.

```{r}

# method 1
y<-mtcars
group_median<-aggregate(cbind(y$mpg, y$disp, y$hp, y$wt), list(y$cyl), median)
colnames(group_median)<-c("cyl", "mpg","disp","hp","wt")
group_median

# method 2
#d<-aggregate(. ~ cyl, mtcars, median)
#d[,c(1,2,3,4,6)]


```



### Exercise 3.19

Test if the variables mpg, disp, hp, wt and qsec have significant difference across the three cyl in the R build data set mtcars. Fill the blanks and interpret your conclusion based on the p-values.

```{r}
aov.fun.car <- function(temx){
m2 <- aov(temx ~ mtcars$cyl)
summary(m2)
}
sub.car <- mtcars[, c(1,3,4,6,7)]
aov.car <- sapply(sub.car, aov.fun.car)
aov.car

# The p-values of the variables mpg, disp, hp, wt and qsec are all close to 0
# Therefore, we reject the null hypothesis.
# the variables mpg, disp, hp, wt and qsec 
# have significant difference across the three cyl.
# We can conclude that mpg among three cyls are not the same;
# disp among three cyls are not same;
# hp among three cyls are not same;
# wt among three cyls are not same;
# qsec among three cyls are not same.

```