---
title: "Homework5"
author: "Li Zhang"
date: "09/23/2021"
output:
  pdf_document: default
  html_document: default
  word_document: default
---
# Exercise 3.12

Let subset.iris <- as.matrix(iris[1:10, 1:4]), Using apply function to calculate the mean of subset.iris by column.
```{r}

subset.iris <- as.matrix(iris[1:10, 1:4])
apply(subset.iris, 2, mean)
```

# Exercise 3.13

Let subset.iris.2 <- as.matrix(iris[, 1:4]), fill blanks in the find.max function defined below to find the maximal value in each column of subset.iris.2.
```{r}

find.max <- function(x){
  y <- function(xcol){
    return(which.max(xcol))
}
  return(apply(x, 2, y))
}


subset.iris.2 <- as.matrix(iris[, 1:4])
find.max(subset.iris.2)
```

# Exercise 3.14

Using the mosaic.plot() function to blur the image of mona_lisa.pgm by adding mosaic to her eyes region. Your output should look like the graph below.

```{r}

library(pixmap)
mona_lisa <- read.pnm("mona_lisa.pgm",cellres = 1)
#plot(mona_lisa)
#str(mona_lisa)
mosaic.plot <- function(image, yrange, xrange){
  length.y <- length(yrange)
  length.x <- length(xrange)
  image2 <- image
  whitenoise <- matrix(nrow = length.y, ncol = length.x, runif(length.y * length.x))
  image2@grey[yrange, xrange] <- whitenoise
  return(image2)
}
mona_lisa.mosaic <- mosaic.plot(mona_lisa, 60:85, 82:148)
plot(mona_lisa.mosaic)






```


# Exercise 3.15

Fill in blanks of the following function so that the eyes region of Mona Lisa is covered by pure color, not mosaic.

Setting the degree=0, 0.5 and 1 respectively, plot the images of Mona Lisa that eyes region are covered by pure colors.
```{r}

mosaic.plot.2 <- function(picture, yrange, xrange, degree){
length.y <- length(yrange)
length.x <- length(xrange)
pic2 <- picture
pic2@grey[yrange,xrange] <- degree
return(pic2)
}
mona_lisa.mosaic <- mosaic.plot.2(mona_lisa, 60:85, 82:148, 0)
plot(mona_lisa.mosaic)

mona_lisa.mosaic <- mosaic.plot.2(mona_lisa, 60:85, 82:148, 0.5)
plot(mona_lisa.mosaic)

mona_lisa.mosaic <- mosaic.plot.2(mona_lisa, 60:85, 82:148, 1)
plot(mona_lisa.mosaic)


```