---
title: "Homework4"
author: "Li Zhang"
date: "09/13/2021"
output:
  word_document: default
  html_document: default
  pdf_document: default
---


## Exercise 3.2

Suppose a vector is defined as x <- c(12, 56, 31, -5, 7).

a. Calculate the mean of all elements in x and assign it to y.
b. Square each element in x and assign the results to a new vector z.
```{r}
x <- c(12, 56, 31, -5, 7)

# The mean of all elements in x
y <- mean(x)
y

# Square each element in x
z <- x^2
z


```

## Exercise 3.3

Use Google to find functions which set and get the current working directory in R, respectively.
```{r}

#setwd(dir) is used to set the current working directory.
setwd("/Users/zhangli/Desktop")


#getwd returns the current working directory.
getwd()


```


## Exercise 3.4

Use Google to find the function which lists all the files in the current working folder in R.
```{r}

#Function that lists all the files in the current working folder in R
#list.files()

#or we can use this function.
#dir()

```


## Exercise 3.5
Compute 1+2+3… +1000 with one line of R code. Hint: examine the example code for sum( ) function in the R help document.

```{r}

sum(1:1000)

```


## Exercise 3.6
Suppose a vector var1 <- c(NA, 334, 566, 319, NA, -307).

Obtain a new vector var2 which removes all NAs from var1.
Using the argument na.rm to calculate the mean of var1. Make sure you ignore all NAs.
```{r}
# Obtain a new vector var2 which removes all NAs from var1
var1 <- c(NA, 334, 566, 319, NA, -307)
var2 <- var1[!is.na(var1)]

# Using the argument na.rm to calculate the mean of var1
mean(var1, na.rm = TRUE)


```


## Exercise 3.7
Using sample selection function randomly select 10 integers from 1 to 100. Create a vector y which satisfies the following conditions: if an selected integer is an even number, then y returns ‘even,’ otherwise y returns ‘odd.’
```{r}

x <-sample(1:100, 10)
y <- ifelse(x%%2 == 0, "even", "odd")
y


```


## Exercise 3.8
Did Mickey catch more fishes than Tom and Jerry combined? Write R code to verify this statement using the fishes vector and return a TRUE or FALSE value.
```{r}
fishes<- c(7, 3, 9)
fishes[3] > sum(fishes[1:2])

```


## Exercise 3.9
Using the name rather than the index in the vector fisher, assign a character ‘Ten’ to Tom.
```{r}
fisher<- c(7, 4, 9) 
names(fisher) <- c("Tom", "Jerry", "Mickey")
fisher["Tom"] <- 'Ten'
fisher

```


## Exercise 3.10
Create a vector representing the prices of groceries, bread $2.5, milk $3.1, jam $5.3, beer $9.1. And create a bar plot to represent this information.
```{r}
groceries <- c(2.5, 3.1, 5.3, 9.1)
names(groceries) <- c("bread", "milk", "jam", "beer")
groceries
barplot(groceries, main = "Groceries")

```



## Exercise 3.11
Create a vector with 21 integers from -10 to 10, and store it in the x variable. Then create a scatter plot of x^2 against x.
```{r}
x <- seq(-10, 10, length = 21)
plot(x, x^2)

```

